A Few Easy Pieces: \\ How to enhance your relationship with your future self.

### 2011-03-15 workflow.Rnw
Took into account Brian Gs comments

\author{Jake Bowers\thanks{Many thanks to Kevin Quinn and Cara Wong for direct help on this document, Mika LaVaque-Manty, Ben Hansen, and Mark Fredrickson for many useful discussions on this topic, and Kieran Healy for inspiration from afar. The source code for this document may be freely downloaded from \url{https://github.com/jwbowers/workflow}.}}

@@

The first author convenes a three way video conference with the other
collaborators while commuting home in his flying car. The group must
go back to the analyses. Which ones? The ones from 2011? Or 2012? Or
2018? Where are the files? The next day, one member of the group finds
some of the files. Now the checks and re-analyses should be
easy. Right? The student-now-faculty member is the same person and
should remember the reason for those bits of code (or at least should
                                                   remember which series of mouse clicks were used to produce the numbers
                                                   for that crucial table --- and those mouse clicks should not be time
                                                   consuming to redo exactly as they were done in 2011, or was it
                                                   2015?). And, of course, the way Microsoft Word/Stata/SPSS/R/LISREL
stores files and the way that your machine reads and writes them is
the same, right --- since Windows and Mac OS X have always existed and
will always continue to exist more or less as they current exist. And
the group knows exactly which bit of code produced which table and
which figure, right? And they wrote their code following Nagler's
Maxims \citet{nagler1995coding} and being following King's Replication
Standard \citet{king1995replication} right?

@@

All tables and graphs should be generated from code and not clicking
(let alone copying and pasting). If I wanted to re-create the figure
you created but with red lines, and only for people in the South, I
should be able to do so with just a few edits to the code.

@@

\begin{figure}[H]
  \begin{center}
<<fig1plot,echo=FALSE,fig=TRUE,width=3,height=3>>=
par(bty="n",xpd=TRUE,pty="s",tcl=-.25)
<<fig1code>>
@
    \caption{Protest activity by income inequality \citep[from]{norris2009data}.}\label{fig:giniprot}
  \end{center}
\end{figure}

@@

If you need to track down something more than a small detail, but yet
the sleuthing is really a footnote to the article, then you can create
another file to explore and report on those questions for the research
group. The writing in such a memo can be more informal, but the norms
of reproducibility ought to be the same as for the main paper. After
all you will be releasing both documents (probably) publicly upon
publication of the article if not before.

@@

In order to know that ``that plot on page 10'' is from a file called
``fig1.pdf'' you need to document it somehow. I can imagine a system
with very disciplined use of a word processor (like Open Office or
Word) --- perhaps using special auxiliary ``listoffigures.txt'' files
or something. In my experience, it is much easier to just use LaTeX
(which can be used in a very user friendly way via LyX).

@@

All files containing commands operating on data must refer to a data
file. A reference to a data file is a line of code the analysis
program will use to operate on (``load''/ ``open'' / ``get'' /
                                ``use'') the data file. One should not have to edit this line on
different computers or platforms in order to execute this command.
Using R, for example, all analysis files should have
load('thedata.rda') or
\Verb+read.csv('http://www.mywebsite.org/Data/thedata.csv')+ or some
equivalent line in them, and 'thedata.csv' should be stored in some
place easy to find (like in the same directory as the file or perhaps
                    in 'Data/thedata.rda').
It never hurts to drop in a comment about where the data is, if in doubt.

Where should you store the data files? A very obvious solution is to
always make sure that the data file used by a command file is in the
same directory as the command file. More elegant solutions (pointed to
                                                            below) require all co-authors to have the same directory structure so
that load('Data/thedata.rda') means the same thing on all computers
used to work on said project.

The principle of modularity \cite{nagler1995coding} suggests that you
separate data cleaning, processing, recoding, and merging from
analysis in different files. So, perhaps your analysis oriented files
will load('cleandata.rda') or something but with a comment in the code
telling the future you (among others) that cleandata.rda was created
from create-cleandata.R which in turn begins with
read.csv('dirtydata.csv') which, from a comment in the file, was
downloaded from
\Verb+http://www.greatfreedata.gov/dirtydata.csv+ on April 1,
2011. Such a data processing file will typically end with something
like \Verb+save('cleandata.rda')+ so that we are doubly certain about
the provenance of the data.

Now, if in 5 years we wonder where 'cleandata.rda' came from, we might
\Verb+ grep cleandata *.R+ to search for occurrences of this file on
our Unix/OS X system. However, if such searching among files is a
burden, and even nicer solution is to maintain a file for each project
called ``MANIFEST.txt'' or ``INDEX.txt'' or ``README.txt'' which lists
the data and command files with brief descriptions of their functions
and relations.

The principle regarding data is to know where the data came from and
what operations were performed on which set of data. I have seen
command files which do not begin with a call to load data and I have
despaired. What is the meaning of \Verb+please-fit(var1 with var2)+ if
I do not know where var1 var2 come from.\footnote{The command
                                                  \Verb+please-fit+ and some other R functions used in this essay come
                                                  from the \Verb+MayIPleaseDoStatistics+ package which emphasizes
                                                  politeness in data analysis. The commands \Verb+please-fit+ and
                                                  \Verb+please-plot+ can be blocked and \Verb+would-you-please-fit+ or
                                                  \Verb+may-I-please-have-a-plot+ can be required using
                                                  \Verb+options(politeness=99)+ }

@@

 Now, if you had a moment to think in between teaching that new class, reading books for the awards committee, reading application files for the admissions committee, staying home with a sick child, and your own current research, you might say to yourself, ``Before I write new code, I should write a test of the code. I should write a little bit of code that let's me know that my double-bootstrap procedure actually does what it is supposed to do.''

@@

Name your files with evocative and descriptive names. Your
collaborators are less likely to call you at midnight asking for help
if your files are named ``inequality-and-protest-analyses.R'' than if your files are
called ``temp9'' or ``supercalifragilisticexpialidocious'' (Note the
                                                            use of the extension .R to tell us that the file contains R
                                                            commands. Use extensions like this as a standard practice to help you
                                                            and your computer get along.)

@@

Now, imagine that the UN never hears of your work, or even that your
article does not spawn debate. Why then would you spend time to
communicate with your future self and others? My own answer to this
question is because I want my work to be credible and useful to myself
and other scholars even if each article does not immediately change
the world.  What I report in my data analyses should have two main
characteristics: (1) the findings of the work should not be a matter
of opinion and (2) other people could reproduce the findings. That is,
what the work represents is not a matter of opinion but is a shared
experience --- and an experience shared without respect to the
identities of others (although requiring some shared technical
training and research resources).
% Such work should make us change how
% we act --- or at least, it ought to stand on stronger epistemological
% ground than other claims about experiences which ought to be shared or
% shareable.

@@

In the end, following these practices and those recommended by X and Y
in this issue allows your computerized analyses of your collections
stuff of to be credible.  Finally, if the UN quibbles with your
analyses, your future self can shoot the archive required to reproduce
your work (in still intelligible plain text, analyzed using commented
code so that folks can translate to whatever system succeeds R, or
since you used R, you can include a copy of R and all of the R
packages you used in your final analyses in 2018 in the archive
itself.) You can say, "Here is everything you need to reproduce my
work." To be extra helpful you can add "Read the README file for
futher instructions." And then you can get on with your life: maybe
the next great idea will occur when your 4 year old asks a wacky
question after stripping and painting her overly cooperative 1 year
old brother purple, or teaching a class, or in a coffee shop, or on a
quiet walk.

\bibliographystyle{asa}
\bibliography{/Users/jwbowers/Documents/BIB/trunk/big}

>>>>

\author{Jake Bowers\thanks{Many thanks to Mark Fredrickson, Brian
    Gaines, Kieran Healy, Kevin Quinn and Cara Wong for direct help on
    this document, Mika LaVaque-Manty, Ben Hansen, and Mark
    Fredrickson for many useful discussions on this topic. The source
    code for this document may be freely downloaded from
    \url{https://github.com/jwbowers/workflow}.}}

@@

The first author convenes a three way video conference with the other
collaborators while commuting home in his flying car. The group must
go back to the analyses. Which ones? The ones from 2011? Or 2012? Or
2018? Where are the files? The next day, one member of the group finds
some of the files. Now the checks and re-analyses should be
easy. Right? The student-now-faculty member is the same person and
should remember the reason for those bits of code (or at least should
                                                   remember which series of mouse clicks were used to produce the numbers
                                                   for that crucial table --- and those mouse clicks should not be time
                                                   consuming to redo exactly as they were done in 2011, or was it
                                                   2015?). And, of course, the way Microsoft Word/Stata/SPSS/R/LISREL
stores files and the way that your machine reads and writes them is
the same, right --- since Windows and Mac OS X have always existed and
will always continue to exist more or less as they currently exist? And
the group knows exactly which bit of code produced which table and
which figure, right? And they wrote their code following Nagler's
Maxims \citet{nagler1995coding} and being following King's Replication
Standard \citet{king1995replication} right?

@@

All tables and graphs should be generated from code, not clicking
(let alone copying and pasting). If I wanted to re-create the figure
you created but with red lines, and only for people in the South, I
should be able to do so with just a few edits to the code.

@@

\begin{figure}[H]
  \begin{center}
<<fig1plot,echo=FALSE,fig=TRUE,width=3,height=3>>=
par(bty="n",xpd=TRUE,pty="s",tcl=-.25)
<<fig1code>>
@
    \caption{Protest activity by income inequality \citep[from][]{norris2009data}.}\label{fig:giniprot}
  \end{center}
\end{figure}

@@

If you need to track down something more than a small detail, but yet
the sleuthing is really a footnote to the article, then you can create
another file to explore and report on those questions for the research
group. The writing in such a memo can be more informal, but the norms
of reproducibility ought to be the same as for the main paper. After
all, you will be releasing both documents (probably) publicly upon
publication of the article if not before.

@@

In order to know that ``that plot on page 10'' is from a file called
``fig1.pdf'' you need to document it somehow. I can imagine a system
with very disciplined use of a word processor (like Open Office or
Word) --- perhaps using special auxiliary ``listoffigures.txt'' files
or something. In my experience, it is much easier just to use LaTeX
(which can be used in a very user friendly way via LyX).

@@

All files containing commands operating on data must refer to a data
file. A reference to a data file is a line of code the analysis
program will use to operate on (``load''/ ``open'' / ``get'' /
                                ``use'') the data file. One should not have to edit this line on
different computers or platforms in order to execute this command.
Using R, for example, all analysis files should have
\Verb+load('thedata.rda')+ or
\Verb+read.csv('http://www.mywebsite.org/Data/thedata.csv')+ or some
equivalent line in them, and `thedata.csv' should be stored in some
place easy to find (like in the same directory as the file or perhaps
                    in \Verb+'Data/thedata.rda'+).
It never hurts to drop in a comment about where the data is, if in doubt.

Where should one store data files? A very obvious solution is
always to make sure that the data file used by a command file is in the
same directory as the command file. More elegant solutions (pointed to
                                                            below) require all co-authors to have the same directory structure so
that \Verb+load('Data/thedata.rda')+ means the same thing on all computers
used to work on said project.

The principle of modularity \cite{nagler1995coding} suggests that you
separate data cleaning, processing, recoding, and merging from
analysis in different files. So, perhaps your analysis oriented files
will \Verb+load('cleandata.rda')+ or something but with a comment in the code
telling the future you (among others) that cleandata.rda was created
from create-cleandata.R which in turn begins with
\Verb+read.csv('dirtydata.csv')+ which, from a comment in the file, was
downloaded from
\Verb+http://www.greatfreedata.gov/dirtydata.csv+ on April 1,
2011. Such a data processing file will typically end with something
like \Verb+save('cleandata.rda')+ so that we are doubly certain about
the provenance of the data.

Now, if in 5 years we wonder where `cleandata.rda' came from, we might
\Verb+ grep cleandata *.R+ to search for occurrences of this file on
our Unix/OS X system. However, if such searching among files is a
burden, an even nicer solution is to maintain a file for each project
called ``MANIFEST.txt'' or ``INDEX.txt'' or ``README.txt'' which lists
the data and command files with brief descriptions of their functions
and relations.

The principle regarding data is to know where the data came from and
what operations were performed on which set of data. I have seen
command files which do not begin with a call to load data, and I have
despaired. What is the meaning of \Verb+please-fit(var1 with var2)+ if
I do not know where var1 var2 come from.\footnote{The command
                                                  \Verb+please-fit+ and some other R functions used in this essay come
                                                  from the \Verb+MayIPleaseDoStatistics+ package which emphasizes
                                                  politeness in data analysis. The commands \Verb+please-fit+ and
                                                  \Verb+please-plot+ can be blocked and \Verb+would-you-please-fit+ or
                                                  \Verb+may-I-please-have-a-plot+ can be required using
                                                  \Verb+options(politeness=99)+ }

@@

 Now, if you had a moment to think in between teaching that new class, reading books for the awards committee, reading application files for the admissions committee, staying home with a sick child, and undertaking the odd bit of your own current research, you might say to yourself, ``Before I write new code, I should write a test of the code. I should write a little bit of code that let's me know that my double-bootstrap procedure actually does what it is supposed to do.''

@@

Name your files with evocative and descriptive names. Your
collaborators are less likely to call you at midnight asking for help
if your files are named ``inequality-and-protest-analyses.R'' than if your files are
called ``temp9'' or ``supercalifragilisticexpialidocious.'' (Note the
                                                            use of the extension .R to tell us that the file contains R
                                                            commands. Use extensions like this as a standard practice to help you
                                                            and your computer get along.)

@@

But what if the UN never hears of your work, or, by some cruel fate, that your
article does not spawn debate? Why then would you spend time to
communicate with your future self and others? My own answer to this
question is that I want my work to be credible and useful to myself
and other scholars even if each article does not immediately change
the world.  What I report in my data analyses should have two main
characteristics: (1) the findings of the work should not be a matter
of opinion; and, (2) other people should be able to reproduce the findings. That is,
what the work represents is not a matter of opinion but is a shared
experience --- and an experience shared without respect to the
identities of others (although requiring some shared technical
training and research resources).
% Such work should make us change how
% we act --- or at least, it ought to stand on stronger epistemological
% ground than other claims about experiences which ought to be shared or
% shareable.

@@

In the end, following these practices and those recommended by X and Y
in this issue allows your computerized analyses of your collections
stuff of to be credible.  Finally, if the UN quibbles with your
analyses, your future self can shoot the archive required to reproduce
your work (in still intelligible plain text, analyzed using commented
code so that folks can translate to whatever system succeeds R, or
since you used R, you can include a copy of R and all of the R
packages you used in your final analyses in 2018 in the archive
itself.) You can say, ``Here is everything you need to reproduce my
work." To be extra helpful you can add ``Read the README file for
futher instructions." And then you can get on with your life: maybe
the next great idea will occur when your 4-year-old asks a wacky
question after stripping and painting her overly cooperative
1-year-old brother purple, or teaching a class, or in a coffee shop,
or on a quiet walk.

\bibliographystyle{apsr}
\bibliography{/Users/jwbowers/Documents/BIB/trunk/big}
